Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-07-23 exp1/iswc2013.tex
add TODO
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@617 43ef1190-90fd-4dd7-a769-d0121f02b80e
Upcase section
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@618 43ef1190-90fd-4dd7-a769-d0121f02b80e

In an experiment, there are several factors, and one may run the
experiment by measuring the response variable for every combination of
levels. In a two-level experiment, this will result in $2^n$ runs,
where $n$ is the number of factors. This is called a \emph{full
  factorial experiment}. This is done in Section~\ref{sec:full}.

@@

It is a fundamental property of two-level experiments that they can
only model linear effects, however, they are usually quite sufficient
for identifying significant effects even though the underlying
response may be non-linear.  As we saw in section~\ref{sec:full}, the
choice of levels may be critical, and if non-linearity is expected,
more than two levels must be considered, or at the very least,
attention must be paid to the choice of levels.

@@

We saw in section~\ref{sec:frac} that with a smaller experiment, fewer
interactions are significant, and so provides us with fewer clues to
assess the soundness of the experiment. This should be a focus of
further research.

@@

As evaluating SPARQL endpoints is inherently difficult a simplistic
experiment such as the one we designed should not hold up to
scrutiny. We set out to demonstrate how an experiment could be set up
using DoE, and show how it can be analyzed. In
sections~\ref{sec:full}~and~\ref{sec:frac}, we first saw how the
analysis correctly pointed out the most important effects, under
assumptions dictated by the factors and levels that were given. We saw
how the formalism provided a comprehensive view of the
experiment. Then, we saw that the formalism pointed out weaknesses
that could invalidate the experiment, and in
section~\ref{sec:hugefrac} we saw that the experiment did indeed not
hold up to scrutiny.

>>>>

In an experiment, there are several factors, and one may run the
experiment by measuring the response variable for every combination of
levels. In a two-level experiment, this will result in $2^n$ runs,
where $n$ is the number of factors. This is called a \emph{full
  factorial experiment}. This is done in Section~\ref{sec:full}.
%TODO {Simple 8 x 2 overview table?}

@@

It is a fundamental property of two-level experiments that they can
only model linear effects, however, they are usually quite sufficient
for identifying significant effects even though the underlying
response may be non-linear.  As we saw in Section~\ref{sec:full}, the
choice of levels may be critical, and if non-linearity is expected,
more than two levels must be considered, or at the very least,
attention must be paid to the choice of levels.

@@

We saw in Section~\ref{sec:frac} that with a smaller experiment, fewer
interactions are significant, and so provides us with fewer clues to
assess the soundness of the experiment. This should be a focus of
further research.

@@

As evaluating SPARQL endpoints is inherently difficult a simplistic
experiment such as the one we designed should not hold up to
scrutiny. We set out to demonstrate how an experiment could be set up
using DoE, and show how it can be analyzed. In
sections~\ref{sec:full}~and~\ref{sec:frac}, we first saw how the
analysis correctly pointed out the most important effects, under
assumptions dictated by the factors and levels that were given. We saw
how the formalism provided a comprehensive view of the
experiment. Then, we saw that the formalism pointed out weaknesses
that could invalidate the experiment, and in
Section~\ref{sec:hugefrac} we saw that the experiment did indeed not
hold up to scrutiny.
